\begin{homeworkProblem}[Monotone matrices (10 points)]

    A $m$-by-$n$ matrix $A$ is \emph{monotone} if $n \geq m$, each row of $A$ 
    has no duplicate entries, and it has the following property: if the minimum 
    of row $i$ is located at column $j_i$, then $j_1 < j_2 < j_3 \ldots j_m$. 
    For example, the following 3-by-6 matrix is monotone (the minimum of each 
    row is bolded), since $j_1 = 1, j_2 = 3, j_3 = 6$:
    
    \[
        \begin{bmatrix}
            \textbf{1} & 3 & 4 & 6 & 5 & 2 \\
            7 & 3 & \textbf{2} & 5 & 6 & 4 \\
            7 & 9 & 6 & 3 & 10 & \textbf{0}
        \end{bmatrix}
    \]

    Give an efficient (i.e., significantly better than $O(mn)$-time) algorithm 
    that finds the minimum in each row of an $m$-by-$n$ monotone matrix $A$.

    Bound the running time of your algorithm. Note: you might find it easier to 
    bound the work per level of recursion directly rather than writing a formal 
    recurrence relation.

    \solution \\
    Below is my algorithm description of a divide-and-conquer algorithm 
    \alg{FindRowMinimums(A)} presented in pseudocode.

    \begin{algorithm}[h]
        \begin{algorithmic}[1]
            \Function{FindRowMinimums}{$A$}
                \If{$A$ has one row}
                    \State{} \Return{minimum element of $A[0]$}
                \EndIf{}
                \State{} $mid \gets \lfloor |A| / 2 \rfloor$
                \State{} Find minimum of row $A[mid]$ and let $minCol$ be its column index
                \State{} $minVal \gets A[mid][minCol]$
                \State{}
                \State{} $top \gets A[0 \ldots mid-1, 0 \ldots minCol-1]$
                \State{} $bottom \gets A[mid+1 \ldots |A|-1, minCol+1 \ldots n-1]$
                \State{}
                \State{} $topMins \gets$ \Call{FindRowMinimums}{$top$}
                \State{} $bottomMins \gets$ \Call{FindRowMinimums}{$bottom$}
                \State{}
                \State{} \Return{$topMins + [minVal] + bottomMins$}
            \EndFunction{}
        \end{algorithmic}
    \end{algorithm}

    The algorithm \alg{FindRowMinimums(A)} produces the correct result on an 
    $m$-by-$n$ monotone matrix because it exploits the monotone property to 
    partition the matrix into independent subproblems. When we find the minimum 
    of the middle row at column $minCol$, the monotone property guarantees that
    all rows above the middle row have their minimums at columns $< minCol$ 
    (since $j_1 < j_2 < \ldots < j_{mid}$). All rows below the middle row have 
    their minimums at columns $> minCol$ (since $j_{mid} < j_{mid+1} < \ldots < 
    j_m$). Therefore, the top subproblem can be restricted to columns $[0, 
    minCol-1]$ and the bottom subproblem to columns $[minCol+1, n-1]$, knowing 
    that no row minimum is excluded. Each recursive call maintains the monotone 
    property on its submatrix, and the base case correctly returns the single 
    row's minimum.

    At each recursion level, at most $n$ entries are scanned in finding the 
    minimum. Additionally, the subproblem is approximately halved each time, 
    therefore leading to $O(\log m)$ levels. Therefore, the runtime of the
    algorithm is $O(n \log m)$. 

\end{homeworkProblem}
