\begin{homeworkProblem}[Shaving Logs or Coding (10 points)]

For full credit, you should do \textbf{one of the following three} questions. (You may solve the others for fun if you want!)

\begin{enumerate}[(a)]
    \item Find a $O(n)$ time algorithm for maximum subarray sum (Question 2), if your solution above was slower.
    
    \item Find a $O(n)$ query algorithm for werewolves (Question 4), if your solution above was slower.
    
    \item Implement the quickselect algorithm in a python jupyter notebook called \texttt{quick\_select.ipynb}. There are two ways that you can access the notebook and complete the problems:
    
    \begin{enumerate}[(a)]
        \item \textbf{On Local Machine}: \texttt{git clone} (or if you already have it, \texttt{git pull}) from the coding homework repo,
        
        \url{https://github.com/Berkeley-CS170/cs170-sp26-coding}
        
        and navigate to the \texttt{hw02} folder. Refer to the \texttt{README.md} for local setup instructions.
        
        \item \textbf{On Datahub}: Click on \url{https://github.com/Berkeley-CS170/cs170-sp26-coding} and navigate to the \texttt{hw02} folder if you prefer to complete this question on Berkeley DataHub.
    \end{enumerate}
    
    Notes:
    \begin{itemize}
        \item \emph{Submission Instructions}: Please run the last cell to download a zip file and submit it to the gradescope assignment titled ``HW02 (Coding)''.
        
        \item \emph{OH/HWP Instructions}: Designated coding course staff will provide conceptual and debugging help during office hours.
        
        \item \emph{Academic Honesty Guideline}: We realize that code for some of the algorithms we ask you to implement may be readily available online, but we strongly encourage you to not directly copy code from these sources. Instead, try to refer to the resources mentioned in the notebook and come up with code yourself. That being said, we \textbf{do acknowledge} that there may not be many different ways to code up particular algorithms and that your solution may be similar to other solutions available online.
    \end{itemize}
\end{enumerate}

\end{homeworkProblem}
