\documentclass[11pt]{article}
\usepackage{../cs170}

\pgfplotsset{compat=1.18}

\def\title{Homework 1}
\def\duedate{Monday 1/26/2026, at 10:00 pm (grace period until 11:59pm)}

\begin{document}
\maketitle
Due \textbf{\duedate}

\question{Study Group}
List the names and SIDs of the members in your study group. If you
have no collaborators, explicitly write ``none''.

\question{Course Policies}

Please read through the course policies below:
\begin{enumerate}
\item The exams will be at the following times. We do not plan on
    offering alternate exams, so please mark your calendars:
    \begin{enumerate}
        \item \textbf{Midterm 1}: Thursday, 2/26/2026, 7:00 PM - 9:00 PM
        \item \textbf{Midterm 2}: Thursday, 4/9/2026, 8:00 PM - 10:00 PM
        \item \textbf{Final}: Friday, 5/15/2026, 7:00 PM - 10:00 PM
    \end{enumerate}

\item Homework is due Mondays at 10:00pm, with a late deadline at
    11:59pm to protect against technical issues. There is no penalty for
    submitting before the late deadline, but absolutely no submissions
    will be accepted after 11:59pm.

\item Your lowest homework will be dropped at the end of the semester.
    By filling out the mid-semester survey and end-of-semester course
    evaluation, you have the opportunity for two additional drops.

\item Each homework will have one question for you to attempt on your
    own, labeled "solo question." You will get detailed feedback on your
    answer to help callibrate you on your learning. On the remaining
    questions you may collaborate with your fellow students and with AI
    tools, but you must always list your collaborators and write up the
    solutions in your own words. (Details in the policies linked below.)

\item The primary source of communication for CS170 will be through
    Edstem.
\item \textbf{Syllabus and Policies:}
    \url{https://cs170.org/policies/}
\item \textbf{Homework Guidelines:}
    \url{https://cs170.org/resources/homework-guidelines/}
\item \textbf{Regrade Etiquette:}
    \url{https://cs170.org/resources/regrade-etiquette/}
\item \textbf{Forum Etiquette:}
    \url{https://cs170.org/resources/ed-etiquette/}
\end{enumerate}
Copy and sign the following sentence on your homework submission.

``I have read and understood the course syllabus and policies.''

\newpage

\question{Asymptotic Complexity Comparisons}

Order the following functions so that for all $i,j$, if $f_i$ comes
before $f_j$ in the order then $f_i = O(f_j)$. Do not justify your
answers.
\begin{itemize}
    \item $f_1(n) = 3^{n}$
    \item $f_2(n) = n^{\frac{1}{3}}$
    \item $f_3(n) = 12$
    \item $f_4(n) = 2^{\log_2 n}$
    \item $f_5(n) = \sqrt{n}$
    \item $f_6(n) = 2^n$
    \item $f_7(n) = \log_2 n$
    \item $f_8(n) = 2^{\sqrt n}$
    \item $f_9(n) = n^3$
    \item $f_{10}(n) = \log_3 n$
    \item $f_{11}(n) = \log^2 n$
\end{itemize}

As an answer you may just write the functions as a list, e.g. $f_8,
f_9, f_1, \ldots$

ß\newpage

\question{Counting Steps (Solo Question)}

You can climb a ladder with $n$ rungs by climbing either 1 rung or 2
rungs in each step. How many distinct ways are there to climb to the
top?
\begin{subparts}
\subpart Give a simple recursive algorithm to compute the answer.

\subpart Prove correctness using induction.

\end{subparts}

\newpage

\question{Counting Steps Efficiently}

\begin{subparts}
\subpart Analyze the running time of your algorithm from the previous
question. Is it bounded by a polynomial in $n$? Assume for simplicity
that two integers can be added in one timestep.

\subpart Give an efficient algorithm to compute the answer, and show
the running time is bounded by $O(n)$ (polynomial in $n$ is also
acceptable).

\subpart (Extra credit) Give an even more efficient algorithm to
compute the answer, and analyze its running time, which should be
sub-polynomial, again assuming each integer arithmetic operation takes
$1$ step. {\it Hint: can you compute the answer by multiplying
$2\times 2$ matrices?}

Is this running time a fair assessment of how the algorithm will
perform in practice?

\end{subparts}

\end{document}
