\begin{homeworkProblem}[Course Policies]
    Please read through the course policies below:
    
    \begin{enumerate}
        \item The exams will be at the following times. We do not plan on 
            offering alternate exams, so please mark your calendars:
            \begin{enumerate}
                \item \textbf{Midterm 1}: Thursday, 2/26/2026, 7:00 PM - 9:00 PM
                \item \textbf{Midterm 2}: Thursday, 4/9/2026, 8:00 PM - 10:00 PM
                \item \textbf{Final}: Friday, 5/15/2026, 7:00 PM - 10:00 PM
        \end{enumerate}

        \item Homework is due Mondays at 10:00pm, with a late deadline at 
            11:59pm to protect against technical issues. There is no penalty 
            for submitting before the late deadline, but absolutely no 
            submissions will be accepted after 11:59pm.

        \item Your lowest homework will be dropped at the end of the semester. 
            By filling out the mid-semester survey and end-of-semester course 
            evaluation, you have the opportunity for two additional drops.

        \item Each homework will have one question for you to attempt on your 
            own, labeled ``solo question." You will get detailed feedback on 
            your answer to help callibrate you on your learning. On the 
            remaining questions you may collaborate with your fellow students 
            and with AI tools, but you must always list your collaborators and 
            write up the solutions in your own words. (Details in the policies 
            linked below.)

        \item The primary source of communication for CS170 will be through 
            Edstem.

        \item \textbf{Syllabus and Policies:} \url{https://cs170.org/policies/}
        \item \textbf{Homework Guidelines:} \url{https://cs170.org/resources/homework-guidelines/}
        \item \textbf{Regrade Etiquette:} \url{https://cs170.org/resources/regrade-etiquette/}
        \item \textbf{Forum Etiquette:} \url{https://cs170.org/resources/ed-etiquette/} 
    \end{enumerate}
    
    Copy and sign the following sentence on your homework submission.

    ``I have read and understood the course syllabus and policies.''

    \begin{homeworkAnswer}
        \textbf{I have read and understood the course syllabus and policies.}

    \end{homeworkAnswer}
    
\end{homeworkProblem}